A large and variable leading tail of Helium in a sub-Saturn undergoing runaway inflation

### 2023-06-05 paper/paper1/ms.tex
3D hydro section is a placeholder
Add an abstract

\begin{abstract}
    To be overhauled...
\end{abstract}

@@

\subsection{3D Hydrodynamic simulations}\label{sec3Dsimulations}
We explore the observational impact of dayside mass loss with 3D hydodynamic models.  The mechanics follow previous such simulations on WASP 107~b \citep{2022ApJ...926..226M} and HAT-P-32~b (Zhang et al. 2023).  We modified the simulations to match the system properties of HAT-P-67~b.  We ran two simulations: one in which the planetary boundary conditions are identical everywhere on the sphere of the planet (the homogeneous or isotropic case). In the
other (asymmetric or anisotropic case) we let the temperature correspond to the incident irradiation through a $\cos(\theta)$ term where $\theta$ is the substellar angle.  Figures \ref{fig:IsotropicMdot} and \ref{fig:DaysideMdot} compare these two scenarios.  The dayside heating case clearly produces a large leading tail comparable to the scale observed in the Helium excess absorption.
\authorcomment1{Insert Morgan Macleod text here.}

>>>>

\begin{abstract}
    Atmospheric escape and its associated mass loss shape the fate of exoplanets, with statistical evidence for photoevaporation imprinted across the mass-radius-insolation distribution. The Habitable Zone Planet Finder (HPF) Spectrograph Exospheres program aims to expand the search across a diverse range of planet and host star properties via the 10830 Ångstrom Helium triplet, a sensitive probe of active atmospheric escape. Here we present a detection of 10\% absorption depth during transit spectroscopy of the highly irradiated, low-gravity, inflated hot Saturn HAT-P-67b. The 13.8 hours of on-sky integration time over 41 HET tracks provides enough out-of-transit phase coverage as to reveal comparably deep absorption preceding the transit, evincing a large leading tail. This configuration can be understood as the escaping material overflowing its small Roche lobe, and advecting most of the gas into the stellar—and not planetary---rest frame, consistent with the Doppler velocity structure seen in the Helium line profiles. We, therefore, attribute the leading tail to a combination of Keplerian shear and stellar wind confinement, with preferential mass loss from dayside heating.  We see some transit-to-transit variability in the line profile, consistent with the interplay of stellar and planetary wind.  We employ 1D Parker wind models and 3D hydrodynamic models to estimate the mass loss rate, finding values on the order of $>10^{13}$ g/s, with large uncertainties owing to the unknown XUV flux of the F host star.  The large mass loss in HAT-P-67b represents a rare example of an inflated hot Saturn, a class of planets predicted by Thorngren and Fortney (2018) to have been statistically depopulated through runaway inflation.  We explore two physical mechanisms for runaway evaporation: Ohmic Dissipation and XUV irradiation, slightly favoring the latter.   
\end{abstract}

@@

\subsection{3D Hydrodynamic simulations}\label{sec3Dsimulations}
We explore the observational impact of dayside mass loss with 3D hydodynamic models.  The mechanics follow previous such simulations on WASP 107~b \citep{2022ApJ...926..226M} and HAT-P-32~b (Zhang et al. 2023).  We modified the simulations to match the system properties of HAT-P-67~b.  We ran two simulations: one in which the planetary boundary conditions are identical everywhere on the sphere of the planet (the homogeneous or isotropic case). In the
other (asymmetric or anisotropic case) we let the temperature correspond to the incident irradiation through a $\cos(\theta)$ term where $\theta$ is the substellar angle.  Figures \ref{fig:IsotropicMdot} and \ref{fig:DaysideMdot} compare these two scenarios.  The dayside heating case clearly produces a large leading tail comparable to the scale observed in the Helium excess absorption.
\authorcomment1{The above paragraph is a placeholder stub slated for revision from Morgan.}
